---
title: "FinalProject"
author: "Nicholas Shuble,Luke Pedersen, Joseph Downs"
date: "2023-11-28"
output: html_document
---
Dataset Introduction:
The dataset we have chosen is from kaggle.com. It is a dataset containing every player in the fifa video game as a case, and attributes such as height, weight, overall rating, etc.




```{r}
library(readr)
# Opening the csv file
fifa_players <- read_csv("archive/fifa_players.csv")
View(fifa_players)

```

```{r}
library(dplyr)
library(ggplot2)
#Selecting the relevant attributes from the dataset as we do not need all of them
fifa_players <- fifa_players  %>%
  select(name, age, height_cm, weight_kgs, positions, nationality, overall_rating, potential, value_euro, wage_euro, preferred_foot, release_clause_euro, crossing, finishing, heading_accuracy, dribbling, curve, acceleration, agility, jumping, stamina)
```

Do taller players have better heading rating?
```{r}
#Create a wrangled data set that only includes necessary values
heights_heading <- fifa_palyers %>% 
  select(height_cm,heading_accuracy)
#Create a scatterplot
ggplot(heights_heading,aes(x = height_cm, y = heading_accuracy)) +
  geom_point(size = 0.7,color = 'red') +
#Create a line of best fit for the scatterplot
  stat_smooth(method = "lm",
        col = "black", se = FALSE, size = 1) +
  labs(title = paste("Height vs. Heading Accuracy")) +
  theme_minimal()
```


Question:
Is there a correlation between Player Potential and Release Clause?

Background: In soccer player potential is fairly self explanatory, it just is a rating that describes what the ceiling of a players ability potentially could be. A release clause is a fee built in to a players contract that allows other clubs to buy out the players contract for said fee without having to negotiate with the players current club, assuming that the player is agreeable. 

Why do we care?:
The reason that this question is interesting, is because if the player has a high potential, then you would assume that their release clause would be higher.


```{r}
fifa_players %>%
 filter(potential >= 59L & potential <= 95L) %>%
 ggplot() +
  aes(x = potential, y = release_clause_euro) +
  geom_line(colour = "#111C46") +
  labs(x = "Player potential", y = "Release Clause (euros)", title = "Potential vs Release Clause") +
  theme_minimal()
```

Summary: based on the results from our data visualization, we can conclude that if a players potential is 70 or higher, their is an exponential rate of increase between potential and release clause, so there is a relationship between these two variables.


Question:



```{r}
fifa_players %>%
 filter(overall_rating >= 60L & overall_rating <= 94L) %>%
 ggplot() +
  aes(x = overall_rating, y = wage_euro) +
  geom_point(shape = "circle", size = 0.5, colour = "#B22222") +
  geom_smooth(span = 0.75) +
  labs(
    x = "Wage (euros)",
    y = "Overall Rating",
    title = "Wage vs Overall Rating"
  ) +
  theme_minimal()
```



```{r pressure, echo=True}
library(ggplot2)

ggplot(fifa_players, aes(x = preferred_foot, y = finishing, fill = preferred_foot)) +
  geom_boxplot() +
  labs(title = "Finishing Ability by Preferred Foot",
       x = "Preferred Foot",
       y = "Finishing Rating") +
  theme_minimal()


library(ggplot2)

data("fifa_players")

library(dplyr)

library(knitr)

library(kableExtra)

summaryTable <- fifa_players %>%

group_by(preferred_foot) %>%

summarise(count=n(),

finishing_min=min(finishing),

finishing_q1=quantile(finishing, probs=.25),

finishing_median=median(finishing),

finishing_q3=quantile(finishing, probs=.75),

finishing_max=max(finishing),

finishing_mean=mean(finishing),

finishing_std=sd(finishing))

formatted_table <- kable(summaryTable, format = "html") %>%

kable_styling(bootstrap_options = "striped")
View
```


